---
title: "Chabaudi data analysis"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Install packages needed
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(CellChat)
library(patchwork)
library(sjmisc)
library(cellhashR)
library(tidyr)
library(PloGO2)
library(clustree)
library(sctransform)
#library(scCustomize)
library(scMayoMap)
library(ggplot2)
library(tibble) 
library(reshape2)
options(stringsAsFactors = FALSE)

```


Loading the data rds files
```{r}
dir = "~/Documents/PhD_Project/Analysis/"

BM_chabaudi <- readRDS(paste0(dir, "BM_chabaudi_query_labelled.rds")) # Chabaudi bone marrows
BM_chabaudi$cell_types <- Idents(BM_chabaudi)

SP_chabaudi <- readRDS(paste0(dir, "SP_chabaudi_query_labelled.rds")) # Chabaudi spleens
SP_chabaudi$cell_types <- Idents(SP_chabaudi)

```


```{r}
DimPlot(SP_chabaudi, label = T) + NoLegend()
DimPlot(BM_chabaudi, label = T) + NoLegend()
```

Use the combined object instead
```{r}
chabaudi_combined <- readRDS(paste0(dir, "all_chabaudi_combined_unlabelled.rds"))
DefaultAssay(chabaudi_combined) <- "integrated"
DimPlot(chabaudi_combined, label = T) + NoLegend()
DimPlot(chabaudi_combined, label = T, split.by = "Tissue") + NoLegend()
```

Rename Idents
```{r}
chabaudi_combined <- RenameIdents(chabaudi_combined, 
          `0` = 'Neutrophils', `1` = "Neutrophils", `2` = "Neutrophils",`3` = "B cells", `4` = "Pro-Erythroblasts", 
          `5` = "Neutrophils", `6` = "Monocytes", `7` = "Erythroblasts", `8` = "Neutrophil Progenitors",
          `9` = "Erythroblasts", `10` = "Macrophages", `11` = "Pre-B cells", `12` = "CD4+ T cells",
          `13` = "NK cells", `14` = "HSPCs", `15` = "Pro-B cells", `16` = "CD8+ T cells",
          `17` = "DCs", `18` = "Macrophages", `19` = "Neutrophils")
```


Visualise and save rds
```{r}
mycols = c("B cells" = "#F28385", "CD4+ T cells" = "#C79ECF", "CD8+ T cells" = "#88D987", "DCs" = "#FED4A6", 
           "Macrophages" = "#6BAED6", "Monocytes" = "#F49C5D", "NK cells" = "#FFFF99", "Neutrophils" = "#F29FC6",
           "Pro-Erythroblasts" = "#76C1BB","Erythroblasts"= "#FF9999", "Neutrophil Progenitors" = "#A4D3E8",
           "Pre-B cells" = "#61B985",  
            "Pro-B cells" = "#AEBD93",
             "HSPCs"= "#C89A76")
           

DimPlot(chabaudi_combined, label = T, pt.size = 0.6, label.size = 5, repel=TRUE, cols = mycols)+ NoLegend() +
  theme(#text = element_text(face = "bold"),
        axis.text.x=element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.text.y=element_text(size=15),
        axis.title.y = element_text(size=15),
        )

DimPlot(chabaudi_combined, label = T, pt.size = 0.6, label.size = 5, repel=TRUE, cols = mycols,  split.by = "Tissue")+ NoLegend() +
  theme(#text = element_text(face = "bold"),
        axis.text.x=element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.text.y=element_text(size=15),
        axis.title.y = element_text(size=15),
        )
```
Editing the cell labels in the spleen
```{r}
chabaudi_combined$cell_type <- Idents(chabaudi_combined)
chabaudi_combined@meta.data <- chabaudi_combined@meta.data %>% 
     mutate(cell_types = case_when(
         Tissue == 'spleen' & cell_type == "Monocytes"  ~ 'Macrophages',
         Tissue == 'spleen' & cell_type == "HSPCs" ~ 'Macrophages',
         Tissue == 'spleen' & cell_type == "Neutrophil Progenitors"  ~ 'Neutrophils',
         Tissue == 'spleen' & cell_type == "Pre-B cells"  ~ 'B cells',
         Tissue == 'spleen' & cell_type == "Pro-B cells"  ~ 'B cells',
         TRUE ~ chabaudi_combined@meta.data$cell_type
     ))

Idents(chabaudi_combined) <- "cell_types"
```


Visualise and save rds
```{r}
DimPlot(chabaudi_combined, label = T, pt.size = 0.6, label.size = 7, repel=TRUE, cols = mycols)+ NoLegend() +
  theme(#text = element_text(face = "bold"),
        axis.text.x=element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.text.y=element_text(size=15),
        axis.title.y = element_text(size=15),
        )

DimPlot(chabaudi_combined, label = T, pt.size = 0.6, label.size = 7, repel=TRUE, cols = mycols,  split.by = "Tissue")+ NoLegend() +
  theme(#text = element_text(face = "bold"),
        axis.text.x=element_text(size=15),
        axis.title.x = element_text(size=15),
        axis.text.y=element_text(size=15),
        axis.title.y = element_text(size=15),
        )
```

```{r}
dir <- "/Users/joycekabagenyi/Documents/PhD_Project/Analysis/"
saveRDS(chabaudi_combined, file = paste0(dir, "all_chabaudi_combined_labelled.rds"))
```

Plot edits
```{r}
# My custom colors
mycols_1 = c("B cells" = "#F28385", "CD4+ T cells" = "#C79ECF", "CD8+ T cells" = "#88D987", "DCs" = "#FED4A6", 
           "Macrophages" = "#6BAED6", "Monocytes" = "#F49C5D", "NK cells" = "#FEE9CF", "Neutrophils" = "#F29FC6",
           "Pro-Erythroblasts" = "#76C1BB","Erythroblasts"= "#EB869D", "Neutrophil Progenitors" = "#A4D3E8",
           "Pre-B cells" = "#61B985",  
            "Pro-B cells" = "#3C8F50",
             "HSPCs"= "#C89A76")

mycols = c("B cells" = "#F28385", "CD4+ T cells" = "#C79ECF", "CD8+ T cells" = "#88D987", "DCs" = "#FED4A6", 
           "Macrophages" = "#6BAED6", "Monocytes" = "#F49C5D", "NK cells" = "#FFFF99", "Neutrophils" = "#F29FC6",
           "Pro-Erythroblasts" = "#76C1BB","Erythroblasts"= "#FF9999", "Neutrophil Progenitors" = "#A4D3E8",
           "Pre-B cells" = "#61B985",  
            "Pro-B cells" = "#AEBD93",
             "HSPCs"= "#C89A76")

#my special theme
jk_theme <- theme(plot.title = element_text(size = 40, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12,angle = -45, hjust = 0),
          axis.text.y = element_text(size = 12)) 
```

Ignoring the erythroblasts and neutrophil in bone marrow
```{r}
## Ignoring the erythroblasts and neutrophil
dta <- subset(chabaudi_combined, subset = cell_type %in% 
                       unique(chabaudi_combined$cell_type)[-c(1,6,9,10)]) # Ignoring Erythroblasts and HSPCs

freq.table.group <- as.data.frame(prop.table(x = table(Idents(subset(dta, subset = Tissue == "bone_marrow")),
                                                       subset(dta, subset = Tissue == "bone_marrow")$TimePoint), 
                                             margin = 2))
ggplot(data = freq.table.group, 
       aes(x=Var2, y = Freq, fill=Var1)) + geom_bar(stat="identity",color="black") +
  labs(x="Infection Time Point", y="Proportion of cells",fill="Cell Types") +
      theme_classic() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, margin = margin(t = 10, b = 10, unit = "pt")),
          axis.text.y = element_text(size = 12,)) +
    scale_fill_manual(values = mycols)+
  scale_x_discrete(labels = c("TP_N1" = "0 dpi", "TP_T2" = "8 dpi", "TP_T3" = "14 dpi", 
                                      "TP_T4" = " 23 dpi","TP_T5" = "56 dpi"))
```


cell types over time
```{r}
freq.table.group <- as.data.frame(prop.table(x = table(subset(chabaudi_combined, subset = Tissue == "bone_marrow")$cell_types,
                                                       subset(chabaudi_combined, subset = Tissue == "bone_marrow")$TimePoint), margin = 2))

ggplot(data=freq.table.group, aes(x=Var2, y = Freq, fill=Var1)) + geom_bar(stat="identity",color="black") +
labs(x="Infection Time Point", y="Proportion of cells",fill="Cell Type") + 
  scale_x_discrete(limits = (levels(freq.table.group$Var2)))+
      theme_classic() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, margin = margin(t = 10, b = 10, unit = "pt")),
          axis.text.y = element_text(size = 12)) +
          scale_fill_manual(values = mycols)+
          scale_x_discrete(labels = c("TP_N1" = "0 dpi", "TP_T2" = "8 dpi", "TP_T3" = "14 dpi", 
                                      "TP_T4" = " 23 dpi","TP_T5" = "56 dpi"))

freq.table.group <- as.data.frame(prop.table(x = table(Idents(subset(chabaudi_combined, subset = Tissue == "spleen")),
                                                       subset(chabaudi_combined, subset = Tissue == "spleen")$TimePoint), margin = 2))

ggplot(data=freq.table.group, aes(x=Var2, y = Freq, fill=Var1)) + geom_bar(stat="identity",color="black") +
labs(x="Infection Time Point", y="Proportion of cells",fill="Cell Type") + 
  scale_x_discrete(limits = (levels(freq.table.group$Var2)))+
      theme_classic() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, margin = margin(t = 10, b = 10, unit = "pt")),
          axis.text.y = element_text(size = 12)) +
          scale_fill_manual(values = mycols)+
          scale_x_discrete(labels = c("TP_N1" = "0 dpi", "TP_T2" = "8 dpi", "TP_T3" = "14 dpi", 
                                      "TP_T4" = " 23 dpi","TP_T5" = "56 dpi"))

## Ignoring the erythroblasts only
dta <- subset(chabaudi_combined, subset = cell_type %in% 
                       unique(chabaudi_combined$cell_type)[-c(1,6,9)]) # Ignoring Erythroblasts and HSPCs

freq.table.group <- as.data.frame(prop.table(x = table(Idents(subset(dta, subset = Tissue == "bone_marrow")),
                                                       subset(dta, subset = Tissue == "bone_marrow")$TimePoint), 
                                             margin = 2))
ggplot(data = freq.table.group, 
       aes(x=Var2, y = Freq, fill=Var1)) + geom_bar(stat="identity",color="black") +
  labs(x="Infection Time Point", y="Proportion of cells",fill="Cell Types") +
      theme_classic() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, margin = margin(t = 10, b = 10, unit = "pt")),
          axis.text.y = element_text(size = 12)) +
          scale_fill_manual(values = mycols) +
          scale_x_discrete(labels = c("TP_N1" = "0 dpi", "TP_T2" = "8 dpi", "TP_T3" = "14 dpi", 
                                      "TP_T4" = " 23 dpi","TP_T5" = "56 dpi"))


freq.table.group <- as.data.frame(prop.table(x = table(Idents(subset(dta, subset = Tissue == "spleen")),
                                                       (subset(dta, subset = Tissue == "spleen"))$TimePoint), margin = 2))
library(ggplot2)

ggplot(data = freq.table.group, aes(x = Var2, y = Freq, fill = Var1)) +
    geom_bar(stat = "identity", color = "black") +
    labs(x = "Infection Time Point", y = "Proportion of cells", fill = "Cell Types") +
    theme_classic() +
    theme(plot.title = element_text(size = 16, face = "bold"),
          axis.title.x = element_text(size = 16),
          axis.title.y = element_text(size = 16),
          legend.title = element_text(size = 14),
          legend.text = element_text(size = 12),
          axis.text.x = element_text(size = 12, angle = 0, hjust = 0.5, margin = margin(t = 10, b = 10, unit = "pt")),
          axis.text.y = element_text(size = 12)) +
          scale_fill_manual(values = mycols)+
          scale_x_discrete(labels = c("TP_N1" = "0 dpi", "TP_T2" = "8 dpi", "TP_T3" = "14 dpi", 
                                      "TP_T4" = " 23 dpi","TP_T5" = "56 dpi"))
```
Cell counts per tissue
```{r}

sp_cells <- as.data.frame(table(Idents(subset(chabaudi_combined, subset = Tissue == "spleen")),
                                                       (subset(chabaudi_combined, subset = Tissue == "spleen"))$TimePoint))

bm_cells <- as.data.frame(table(Idents(subset(chabaudi_combined, subset = Tissue == "bone_marrow")),
                                                       (subset(chabaudi_combined, subset = Tissue == "bone_marrow"))$TimePoint))
# Rename variables
 names(bm_cells) <- c("cell_type", "time_point", "bm_counts")
 names(sp_cells) <- c("cell_type", "time_point", "sp_counts")
 
 # Rotate the table to a wider format
 bm_cells <- pivot_wider(bm_cells, names_from = time_point, values_from = bm_counts)
 sp_cells <- pivot_wider(sp_cells, names_from = time_point, values_from = sp_counts)
 
 # Calculate sum for each column
summary_row_bm <- summarise_all(bm_cells[,-1], sum, na.rm = TRUE)

summary_row_sp <- summarise_all(sp_cells[,-1], sum, na.rm = TRUE)

# Combine the bm and sp conts, then sum to get grand total cells time point the summary row to the original dataframe
all_cells <- rbind(summary_row_bm, summary_row_sp)
summary_row <- summarise_all(all_cells, sum, na.rm = TRUE) # Calculate the sums
all_cells <- bind_rows(all_cells, summary_row) # Add the summary row to the original dataframe

#df <- filter(!apply(df, 1, function(x) any(grepl("H2.", x))))})
#df <- filter(!apply(df, 1, function(x) any(grepl("H2.", x))))})

 
# Adding a new variable and positioning it at the start
all_cells <- all_cells %>%
  dplyr::mutate(Tissue = c("Bone marrow", "Spleen", "Total")) %>%  # Adding new column
  dplyr::select(Tissue, everything())           # Reordering to make it first

# Save csv
write.csv(all_cells, "/Users/joycekabagenyi/Documents/PhD_Project/Conferences and workshops/BIOMAL_2024/plots/cell_count.csv")

```


Marker genes plotting with dot plots
```{r}
DefaultAssay(chabaudi_combined) <- "RNA"

custom_order <- c("CD8+ T cells", "CD4+ T cells", "NK cells", "DCs", "Monocytes", "Macrophages", "HSPCs", 
                  "Pro-B cells", "Pre-B cells", "B cells", "Neutrophil Progenitors", "Neutrophils", 
                  "Pro-Erythroblasts", "Erythroblasts"
                  )

# With unlabelled object
VlnPlot(chabaudi_combined, features = c("Cd3e", "Cd8a", "Cd4","Ccl5", "Siglech", "Ctsg", "S100a4",
                                        "Gata2", "Cd79a", "Vpreb1", "Vpreb3", "Bank1", "Ngp", "S100a9", "Ly6g",     
                                         "Hemgn", "Bpgm"))

FeaturePlot(chabaudi_combined, features = c("Cd3e", "Cd8a", "Cd4","Ccl5", "Siglech", "Ctsg", "S100a4",
                                        "Gata2", "Cd79a", "Vpreb1", "Vpreb3", "Bank1", "Ngp", "S100a9", "Ly6g",     
                                         "Hemgn", "Bpgm"), min.cutoff = "q9")

DotPlot(chabaudi_combined, features = c("Cd3e", "Cd8a", "Cd4","Ccl5", "Siglech", "Ctsg", "S100a4",
                                        "Gata2", "Cd79a", "Vpreb1", "Vpreb3", "Bank1", "Ngp", "S100a9", "Ly6g",     
                                         "Hemgn", "Bpgm"),
        cols = c("#F8F6ED", "#A11213")) + 
  scale_y_discrete(limits = rev(custom_order), labels = rev(custom_order)) +
  theme(axis.text.y = element_text(#angle = 45, 
                                   hjust = 1, size = 14)) +
  ylab("Cell types") +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14))+
  theme(axis.title.x = element_text(size = 16)) 

```





Cell Chat

Loading the necessary libraries
```{r}
library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)
library(dplyr)
library(Seurat)

```

Cell chat object
```{r}
cells.present = as.character(unique(chabaudi_combined$cell_types)[-c(1,7,13)]) # select cells excluding erythroblasts and HSPCs
BM_anchors.combined <- subset(chabaudi_combined, subset = cell_types %in% cells.present)            
data.input <- GetAssayData(BM_anchors.combined, assay = "RNA", slot = "data") # normalized data matrix

# Met data
labels <- Idents(BM_anchors.combined)
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels
names(meta)[names(meta) == "group"] <- "labels"

#Getting the extra variables for the meta data

extra.meta = BM_anchors.combined@meta.data
meta = dplyr::bind_cols(meta, extra.meta)

#cellChat <- createCellChat(object = BM_anchors.combined, group.by = "ident", assay = "RNA")
```


Before creating the cell chat object
```{r}
# meta_select = subset(meta, labels != "Late Pro-Bcells")
cell.use1 = rownames(meta)[meta$TimePoint == "TP_N1" & meta$Tissue == "spleen" & meta$cell_types != "DCs"] # cells from spleen naive  
cell.use2 = rownames(meta)[meta$TimePoint == "TP_T2" & meta$Tissue == "spleen" & meta$cell_types != "DCs"] # cells from spleen TP2
cell.use3 = rownames(meta)[meta$TimePoint == "TP_T3" & meta$Tissue == "spleen" & meta$cell_types != "DCs"] # cells from spleen TP3 
cell.use4 = rownames(meta)[meta$TimePoint == "TP_T4" & meta$Tissue == "spleen" & meta$cell_types != "DCs"] # cells from spleen TP4
cell.use5 = rownames(meta)[meta$TimePoint == "TP_T5" & meta$Tissue == "spleen" & meta$cell_types != "DCs"] # cells from spleen TP2
                           
cell.use6 = rownames(meta)[meta$TimePoint == "TP_N1" & meta$Tissue == "bone_marrow" ] # extract the cell names from bone marrow naive  
cell.use7 = rownames(meta)[meta$TimePoint == "TP_T2" & meta$Tissue == "bone_marrow" ] # extract the cell names from bone marrow TP2
cell.use8 = rownames(meta)[meta$TimePoint == "TP_T3" & meta$Tissue == "bone_marrow" ] # extract the cell names from bone marrow TP3  
cell.use9 = rownames(meta)[meta$TimePoint == "TP_T4" & meta$Tissue == "bone_marrow" ] # extract the cell names from bone marrow TP4
cell.use10 = rownames(meta)[meta$TimePoint == "TP_T5" & meta$Tissue == "bone_marrow" ] # extract the cell names from bone marrow TP5 


# Prepare input data for CellChat analysis
cc.obj <- tibble(organ = c(rep("spleen",5), rep("bone_marrow",5)), 
                     timepoint = c(rep(unique(meta$TimePoint),2))) #cell_chat_objects
cc.obj.list <- c()
  
for (i in c(1:10)) {
  cell_file <- paste0("cell.use", i)
  input_file <- paste0("data.input", i)
  chat_file <- paste(cc.obj[i,1], cc.obj[i,2], sep = "_")
  
  #data.input1 <- GetAssayData(subset(BM_anchors.combined, Tissue == "spleen"), assay = "RNA", slot = "data")
  
  # Subset the expression matrix and meta data
  my.data.input = data.input[, get(cell_file)]
  my.meta = meta[get(cell_file), ]
  my.meta$labels = droplevels(my.meta$labels, exclude = setdiff(levels(my.meta$labels),unique(my.meta$labels)))
  
  # Create the cell chat object
  my.cellchat <- createCellChat(object = my.data.input)
  my.cellchat <- addMeta(my.cellchat, meta = my.meta)
  my.cellchat <- setIdent(my.cellchat, ident.use = "labels") # set "labels" as default cell identity
  levels(my.cellchat@idents) # show factor levels of the cell labels
  groupSize <- as.numeric(table(my.cellchat@idents)) # number of cells in each cell group
  
  # Save the cell chat object under the right name
  assign(chat_file, my.cellchat)
  
  cc.obj.list <- append(cc.obj.list, chat_file)

}
```

```{r}
# 1. Set the ligand-receptor interaction database

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)

#Deleting the two unfound interactions from the  "subset" function in previous code and then runing the upper chunk again

which(CellChatDB[["interaction"]]$ligand == "H2-BI") # 1887
CellChatDB[["interaction"]] <- CellChatDB[["interaction"]][-1887,]
which(CellChatDB[["interaction"]]$ligand == "H2-Ea-ps") #1900
CellChatDB[["interaction"]] <- CellChatDB[["interaction"]][-1900,]

# use all CellChatDB for cell-cell communication analysis
CellChatDB.use <- CellChatDB # simply use the default CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use
```



```{r}
for (i in cc.obj.list) {
  
cellchat = get(i)

# 1. Set the ligand-receptor interaction database

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)

#Deleting the two unfound interactions from the  "subset" function in previous code and then runing the upper chunk again

which(CellChatDB[["interaction"]]$ligand == "H2-BI") # 1887
CellChatDB[["interaction"]] <- CellChatDB[["interaction"]][-1887,]
which(CellChatDB[["interaction"]]$ligand == "H2-Ea-ps") #1900
CellChatDB[["interaction"]] <- CellChatDB[["interaction"]][-1900,]

# use all CellChatDB for cell-cell communication analysis
CellChatDB.use <- CellChatDB # simply use the default CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use


# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
#future::plan("multisession", workers = 4) # do parallel

#Part I: Identify over expressed genes and interactions
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

#Part II: Inference of cell-cell communication network
cellchat <- computeCommunProb(cellchat, population.size = TRUE)
#cellchat <- computeCommunProb(cellchat)
cellchat <- filterCommunication(cellchat, min.cells = 10) # Filter out the cell-cell communication if there are only few number of cells in certain cell groups

#Infer the cell-cell communication at a signaling pathway level
cellchat <- computeCommunProbPathway(cellchat)

#Calculate the aggregated cell-cell communication network
cellchat <- aggregateNet(cellchat)

cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network

cellchat <- computeCommunProbPathway(cellchat)

#cellchat <- updateCellChat(cellchat)

assign(i,cellchat)

}

```



Additional analysis on individual cell chat objects
```{r}
for (i in cc.obj.list){
  cellchat = get(i)
  netAnalysis_signalingRole_network(cellchat, width = 8, height = 2.5, font.size = 10) 
  
  cellchat.Pb_Infected <- computeNetSimilarity(cellchat.Pb_Infected, type = "functional")
  cellchat.Pb_Infected <- netEmbedding(cellchat.Pb_Infected, type = "functional")
  cellchat.Pb_Infected <- netClustering(cellchat.Pb_Infected, type = "functional", do.parallel = F)
  
  cellchat.Pc_Infected <- computeNetSimilarity(cellchat.Pc_Infected, type = "functional")
  cellchat.Pc_Infected <- netEmbedding(cellchat.Pc_Infected, type = "functional")
  cellchat.Pc_Infected <- netClustering(cellchat.Pc_Infected, type = "functional",do.parallel = F)
  
  # Structural analysis
  cellchat.Pb_Infected <- computeNetSimilarity(cellchat.Pb_Infected, type = "structural")
  cellchat.Pb_Infected <- netEmbedding(cellchat.Pb_Infected, type = "structural")
  cellchat.Pb_Infected <- netClustering(cellchat.Pb_Infected, type = "structural",do.parallel = F)
  
  cellchat.Pc_Infected <- computeNetSimilarity(cellchat.Pc_Infected, type = "structural")
  cellchat.Pc_Infected <- netEmbedding(cellchat.Pc_Infected, type = "structural")
  cellchat.Pc_Infected <- netClustering(cellchat.Pc_Infected, type = "structural", do.parallel = F)
  
  nPatterns = 4 # for control
  cellchat.Pb_Infected <- identifyCommunicationPatterns(cellchat.Pb_Infected, pattern = "outgoing", k = nPatterns)
  
  nPatterns =  3# for control
  cellchat.Pb_Infected <- identifyCommunicationPatterns(cellchat.Pb_Infected, pattern = "incoming", k = nPatterns)


}
 
```

Merge cell chat objects

```{r}
#comparison_files <- c("TP_T2", "TP_T3", "TP_T4", "TP_T5")
  #group.new = levels(spleen_TP_N1@idents)
  #spleen_TP_T2 <- liftCellChat(spleen_TP_T2, group.new)
  #spleen_TP_T2 <- updateCellChat(spleen_TP_T2)

#Spleens

  object.list <- list(TP_N1 = spleen_TP_N1, TP_T2 = spleen_TP_T2)
  S2 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_N1 = spleen_TP_N1, TP_T3 = spleen_TP_T3)
  S3 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_N1 = spleen_TP_N1, TP_T4 = spleen_TP_T4)
  S4 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_N1 = spleen_TP_N1, TP_T5 = spleen_TP_T5)
  S5 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_T2 = spleen_TP_T2, TP_T3 = spleen_TP_T3)
  S2_S3 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_T2 = spleen_TP_T2, TP_T4 = spleen_TP_T4)
  S2_S4 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  
  
#Bone marrows
  object.list <- list(TP_N1 = bone_marrow_TP_N1, TP_T2 = bone_marrow_TP_T2)
  B2 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_N1 = bone_marrow_TP_N1, TP_T3 = bone_marrow_TP_T3)
  B3 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)

    object.list <- list(TP_N1 = bone_marrow_TP_N1, TP_T4 = bone_marrow_TP_T4)
  B4 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_N1 = bone_marrow_TP_N1, TP_T5 = bone_marrow_TP_T5)
  B5 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_T2 = bone_marrow_TP_T2, TP_T4 = bone_marrow_TP_T4)
  B2_B4 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  
  object.list <- list(TP_T2 = bone_marrow_TP_T2, TP_T3 = bone_marrow_TP_T3)
  B2_B3 <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)


merged_chat_list <- c("S2","S3","S4", "S5","S2_S3", "S2_S4", "B2", "B3", "B4", "B5", "B2_B3", "B2_B4" )
```


```{r}
for (item in c("B2", "B3", "B4", "B5", "B2_B3", "B2_B4" )) {
  cellchat <- get(i)
  levels(cellchat@meta$labels)[levels(cellchat@meta$labels) =='Pre-B cells' ] <- 'B cells' 
  levels(cellchat@meta$labels)[levels(cellchat@meta$labels) =='Pro-B cells' ] <- 'B cells'
  levels(cellchat@meta$labels)[levels(cellchat@meta$labels) =='Neutrophil Progenitors' ] <- 'Pro-Neutrophils'

  cellchat@meta$cell_types <- cellchat@meta[["labels"]]
  cellchat <- setIdent(cellchat, ident.use = "cell_types") # set "cell_type" as default cell identity
  
  assign(item, cellchat)
}

cell_df <- c("Neutrophils", "Monocytes", "DCs", "Macrophages", "Neutrophil Progenitors", "NK cells", "B cells", "CD8+ T cells", "CD4+ T cells" )
```



```{r}
for (i in merged_chat_list) {
  
  cellchat <- get(i)
  
  #ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing")
  #ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming")
  #print(ht1 + ht2)
  
  ptm = Sys.time()
  gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2)) + theme(axis.text.x=element_text(size=12, hjust=1 ),
                                                                             axis.text.y=element_text(size=12),
                                                                             axis.title.y = element_text(size=15),)+ coord_flip() +
  gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight") + theme(axis.text.x=element_text(size=12,  hjust=1 ),
                                                                             axis.text.y=element_text(size=12),
                                                                             axis.title.y = element_text(size=15),)+ coord_flip() +
  print(gg1 + gg2)
  
  netVisual_diffInteraction(cellchat, weight.scale = T) + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),
                                                                             axis.text.y=element_text(size=12))
                                                                             #axis.title.y = element_text(size=15),
  netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight") + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),
                                                                             axis.text.y=element_text(size=12))
                                                                             #axis.title.y = element_text(size=15),)
  
  gg1 <- netVisual_heatmap(cellchat, font.size = 15, font.size.title = 15) 
  #> Do heatmap based on a merged object
  gg2 <- netVisual_heatmap(cellchat, measure = "weight", font.size = 15, font.size.title = 15) 
  #> Do heatmap based on a merged object
  print(gg1)
  print(gg2)
  
  gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE) + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),axis.text.y=element_text(size=12),legend.text=element_text(size=12)) 
  gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE) + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),axis.text.y=element_text(size=12), legend.text=element_text(size=12))
  
  print(gg1 + gg2)
  
  }


```





```{r}
### Identify signals contributing most to outgoing or incoming signaling of certain cell groups
We can also answer the question on which signals contributing most to outgoing or incoming signaling of certain cell groups.
```{r, fig.width=11, fig.height = 6, fig.align = "center"}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming")
ht1 + ht2

rankNet(S2_S4, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE) + rankNet(S2_S4, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE)
```


```{r}
for (i in merged_chat_list) {
    
    cellchat <- get(i)
    
    #ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing")
    #ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming")
    #print(ht1 + ht2)
    
    ptm = Sys.time()
    gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2)) + theme(axis.text.x=element_text(size=12, hjust=1 ),
                                                                                  axis.text.y=element_text(size=12),
                                                                                  axis.title.y = element_text(size=15),)+ coord_flip() 
    gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight") + theme(axis.text.x=element_text(size=12,  hjust=1 ), axis.text.y=element_text(size=12), axis.title.y = element_text(size=15),)+ coord_flip()  
    
    print(gg1)
    }
```









```{r}
data.input1 = data.input[, cell.use1]
meta1 = meta[cell.use1, ]

############### Before processing the control cellchat object, run code here below
#meta1$seurat_clusters <- droplevels(meta1$seurat_clusters, exclude = c("Late Pro-Bcells"))
#x = subset(meta, !labels == "Late Pro-Bcells")
#meta1$labels = droplevels(meta1$labels, exclude = setdiff(levels(meta1$labels),unique(meta1$labels))) #Tofix the earlier encountered error when running the ComputeCellComProbab command

data.input2 = data.input[, cell.use2]
meta2 = meta[cell.use2, ]



cellchat2 <- createCellChat(object = data.input2)
cellchat2 <- addMeta(cellchat2, meta = meta2)
cellchat2 <- setIdent(cellchat2, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat2@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat2@idents)) # number of cells in each cell group

```





Spleen vs Bone marrow head on comparisons
```{r}
  group.new = levels(bone_marrow_TP_T2@idents)
  spleen_TP_T2 <- liftCellChat(spleen_TP_T2, group.new)
  spleen_TP_T2 <- updateCellChat(spleen_TP_T2)
  object.list <- list(Spleen_peak = spleen_TP_T2, Bone_marrow_peak = bone_marrow_TP_T2)
  TP2_comparisons <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
  

  cellchat = TP2_comparisons
  
 gg1 <- netVisual_heatmap(cellchat, font.size = 15, font.size.title = 15) 
  #> Do heatmap based on a merged object
  gg2 <- netVisual_heatmap(cellchat, measure = "weight", font.size = 15, font.size.title = 15) 
  #> Do heatmap based on a merged object
  print(gg1)
  print(gg2)
  

ptm = Sys.time()
  gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2)) + theme(axis.text.x=element_text(size=12, hjust=1 ),
                                                                             axis.text.y=element_text(size=12),
                                                                             axis.title.y = element_text(size=15),)
  gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight") + theme(axis.text.x=element_text(size=12,  hjust=1 ),
                                                                             axis.text.y=element_text(size=12),
                                                                             axis.title.y = element_text(size=15),)
  print(gg1 + gg2)
    
gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE) + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),axis.text.y=element_text(size=12),legend.text=element_text(size=12)) 

gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE) + theme(axis.text.x=element_text(size=12, angle=45, hjust=1 ),axis.text.y=element_text(size=12), legend.text=element_text(size=12))
  
print(gg1)
print(gg2)
```







