---
title: "Multi-NicheNet analysis BioMalPar"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}
# install.packages("devtools")
devtools::install_github("saeyslab/nichenetr")
devtools::install_github("saeyslab/multinichenetr")

```


```{r}
rm(list=ls())
```


```{r}
library(SingleCellExperiment)
library(dplyr)
library(ggplot2)
library(multinichenetr)
```
Import data rds file
```{r}
chabaudi_combined <- readRDS("~/Documents/PhD_Project/Analysis/all_chabaudi_combined_labelled.rds")
DefaultAssay(chabaudi_combined)
unique(Idents(chabaudi_combined))
DimPlot(chabaudi_combined, label = T) + NoLegend()
```
The reference dataset
```{r}
organism = "mouse"
if(organism == "human"){
  lr_network = readRDS(url("https://zenodo.org/record/7074291/files/lr_network_human_21122021.rds"))
  lr_network = lr_network %>% dplyr::rename(ligand = from, receptor = to) %>% distinct(ligand, receptor) %>% mutate(ligand = make.names(ligand), receptor = make.names(receptor))
  ligand_target_matrix = readRDS("/Users/joycekabagenyi/ligand_target_matrix_nsga2r_final.rds")
  colnames(ligand_target_matrix) = colnames(ligand_target_matrix) %>% make.names()
  rownames(ligand_target_matrix) = rownames(ligand_target_matrix) %>% make.names()
} else if(organism == "mouse"){
  lr_network = readRDS(url("https://zenodo.org/record/7074291/files/lr_network_mouse_21122021.rds"))
  lr_network = lr_network %>% dplyr::rename(ligand = from, receptor = to) %>% distinct(ligand, receptor) %>% mutate(ligand = make.names(ligand), receptor = make.names(receptor))
  ligand_target_matrix = readRDS("/Users/joycekabagenyi/ligand_target_matrix_nsga2r_final_mouse.rds")
  colnames(ligand_target_matrix) = colnames(ligand_target_matrix) %>% make.names()
  rownames(ligand_target_matrix) = rownames(ligand_target_matrix) %>% make.names()
}
```


Convert my seurat object into a singleR object # Starting with the bone marrow then spleen
```{r}
BM_chabaudi.combined <- subset(chabaudi_combined, Tissue == "bone_marrow") # Subset seurat object
BM_chabaudi.combined@meta.data[["consensuscall"]] <- ifelse(is.na(BM_chabaudi.combined@meta.data[["consensuscall"]]) == TRUE, BM_chabaudi.combined@meta.data[["Condition"]],BM_chabaudi.combined@meta.data[["consensuscall"]])
unique(BM_chabaudi.combined@meta.data[["consensuscall"]])

BM_chabaudi.combined$sampleID <- (ifelse(is.na(BM_chabaudi.combined@meta.data[["TimePoint"]]) == FALSE,
                                    paste(BM_chabaudi.combined@meta.data[["consensuscall"]],BM_chabaudi.combined@meta.data[["TimePoint"]], sep = "_"),BM_chabaudi.combined@meta.data[["Condition"]]))

# Merge pre- and pro- B cells to B cells and Pro-Neutrophils to Neutrophils
BM_chabaudi.combined <- RenameIdents(BM_chabaudi.combined, 
          "Pro-B cells" = "B cells", "Pre-B cells" = "B cells","Neutrophil Progenitors" = "Neutrophils")
BM_chabaudi.combined$cell_types <- Idents(BM_chabaudi.combined)

sce_chabaudi = Seurat::as.SingleCellExperiment(BM_chabaudi.combined, assay = "RNA")
sce_chabaudi = alias_to_symbol_SCE(sce_chabaudi, "mouse") %>% makenames_SCE()
```

```{r}
sample_id = "sampleID"
group_id = "TimePoint"
celltype_id = "cell_types"
covariates = NA
batches = NA
```



```{r}
SummarizedExperiment::colData(sce_chabaudi)$sampleID = SummarizedExperiment::colData(sce_chabaudi)$sampleID %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$TimePoint = SummarizedExperiment::colData(sce_chabaudi)$TimePoint %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$cell_types = SummarizedExperiment::colData(sce_chabaudi)$cell_types %>% make.names()

unique(colData(sce_chabaudi)$sampleID)
unique(colData(sce_chabaudi)$TimePoint)
unique(colData(sce_chabaudi)$cell_types)

```

```{r}
senders_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
senders_oi = senders_oi[-c(2,6,10)] # Ignore erythroblasts and HSPCs
receivers_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
receivers_oi = receivers_oi[-c(2,6,10)] # Ignore erythroblasts and HSPCs
```

Subset or obeject if we want to use just a set of our cells
```{r}
sce_chabaudi = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %in% c(senders_oi, receivers_oi)] 
# cell type abundance and expression information from receiver and sender cell types
abundance_expression_info = get_abundance_expression_info(sce = sce_chabaudi, sample_id = "sampleID", group_id = "TimePoint", celltype_id = "cell_types", min_cells = min_cells, senders_oi = senders_oi, receivers_oi = receivers_oi, lr_network = lr_network, batches = batches)

abundance_expression_info$abund_plot_sample
abundance_expression_info$abund_plot_group
```
DE
```{r}
DE_info = get_DE_info(sce = sce_chabaudi, sample_id = "sampleID", group_id = "TimePoint", celltype_id = "cell_types", batches = batches, covariates = covariates, contrasts_oi = contrasts_oi, min_cells = min_cells)

DE_info$hist_pvals
```











---------TWO TIMEPOINT COMPARIAONS

```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'TP_N1-TP_T2','TP_T2-TP_N1'")
contrast_tbl = tibble(
  contrast = c("TP_N1-TP_T2","TP_T2-TP_N1"), 
  group = c("TP_N1","TP_T2"))
```

We subset the data to keep only cells from our timepoints that we want to compare first
```{r}
sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.3
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE
```

For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
```{r}
top_n_target = 250
# Then set multicore analysis to process faster
cores_system = 8
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type
```

Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
```{r}
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)
```


Perform MultiNicheNet’s cell-cell communication analysis
```{r}
multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```
Save output
```{r}
BM_TPN1_vs_TP2 <- multinichenet_output
BM_TPN1_vs_TP2_merged <- multinichenet_output
```

NOW TP_N1 AND TP_TP4

```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'TP_N1-TP_T4','TP_T4-TP_N1'")
contrast_tbl = tibble(
  contrast = c("TP_N1-TP_T4","TP_T4-TP_N1"), 
  group = c("TP_N1","TP_T4"))

```

We subset the data to keep only cells from our timepoints that we want to compare first
```{r}
sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.05
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE
```

For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
```{r}
top_n_target = 250
# Then set multicore analysis to process faster
cores_system = 8
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type
```

Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
```{r}
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)
```


Perform MultiNicheNet’s cell-cell communication analysis
```{r}
multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```
Save output
```{r}
BM_TPN1_vs_TP4 <- multinichenet_output
BM_TPN1_vs_TP4_merged <- multinichenet_output
```






**********************SPLEEN

Convert my seurat object into a singleR object # Starting with the bone marrow then spleen
```{r}
SP_chabaudi.combined <- subset(chabaudi_combined, Tissue == "spleen") # Import RDS
SP_chabaudi.combined@meta.data[["consensuscall"]] <- ifelse(is.na(SP_chabaudi.combined@meta.data[["consensuscall"]]) == TRUE, SP_chabaudi.combined@meta.data[["Condition"]],SP_chabaudi.combined@meta.data[["consensuscall"]])
unique(SP_chabaudi.combined@meta.data[["consensuscall"]])

SP_chabaudi.combined$sampleID <- (ifelse(is.na(SP_chabaudi.combined@meta.data[["TimePoint"]]) == FALSE,
                                    paste(SP_chabaudi.combined@meta.data[["consensuscall"]],SP_chabaudi.combined@meta.data[["TimePoint"]], sep = "_"),SP_chabaudi.combined@meta.data[["Condition"]]))
#BM_chabaudi.combined
sce_chabaudi = Seurat::as.SingleCellExperiment(SP_chabaudi.combined, assay = "RNA")
sce_chabaudi = alias_to_symbol_SCE(sce_chabaudi, "mouse") %>% makenames_SCE()
```

```{r}
sample_id = "sampleID"
group_id = "TimePoint"
celltype_id = "cell_types"
covariates = NA
batches = NA
```


```{r}
SummarizedExperiment::colData(sce_chabaudi)$sampleID = SummarizedExperiment::colData(sce_chabaudi)$sampleID %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$TimePoint = SummarizedExperiment::colData(sce_chabaudi)$TimePoint %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$cell_types = SummarizedExperiment::colData(sce_chabaudi)$cell_types %>% make.names()

unique(colData(sce_chabaudi)$sampleID)
unique(colData(sce_chabaudi)$TimePoint)
unique(colData(sce_chabaudi)$cell_types)

```

```{r}
senders_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
senders_oi = senders_oi[-c(1,7)] # Ignore erythroblasts and HSPCs
receivers_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
receivers_oi = receivers_oi[-c(1,7)] # Ignore erythroblasts and HSPCs
```

Subset or obeject if we want to use just a set of our cells
```{r}
sce_chabaudi = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %in% c(senders_oi, receivers_oi)]

# cell type abundance and expression information from receiver and sender cell types
abundance_expression_info = get_abundance_expression_info(sce = sce_chabaudi, sample_id = "sampleID", group_id = "TimePoint", celltype_id = "cell_types", min_cells = min_cells, senders_oi = senders_oi, receivers_oi = receivers_oi, lr_network = lr_network, batches = batches)

abundance_expression_info$abund_plot_sample
abundance_expression_info$abund_plot_group
```


DE
```{r}
DE_info = get_DE_info(sce = sce_chabaudi, sample_id = "sampleID", group_id = "TimePoint", celltype_id = "cell_types", batches = batches, covariates = covariates, contrasts_oi = contrasts_oi, min_cells = min_cells)

DE_info$hist_pvals
```


_______THREE TIMEPOINTS COMPARISONS

Bone Marrow

```{r}
contrasts_oi = c("'TP_N1-(TP_T2+TP_T4)/2','TP_T2-(TP_N1+TP_T4)/2','TP_T4-(TP_T2+TP_N1)/2'")
contrast_tbl = tibble(contrast = 
                        c("TP_N1-(TP_T2+TP_T4)/2","TP_T2-(TP_N1+TP_T4)/2", "TP_T4-(TP_T2+TP_N1)/2"), 
                      group = c("TP_N1","TP_T2","TP_T4"))

# Perform DE for each cell type
DE_info = get_DE_info(sce = sce, sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, batches = batches, covariates = covariates, contrasts_oi = contrasts_oi, min_cells = min_cells)

DE_info$hist_pvals # View DE out puts

sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.05
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE

# For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
top_n_target = 250
cores_system = 8 # Then set multicore analysis to process faster
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type



# Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)

# Perform MultiNicheNet’s cell-cell communication analysis

multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```




----- TWO TIMEPOINT COMPARISONS

```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'TP_N1-TP_T2','TP_T2-TP_N1'")
contrast_tbl = tibble(
  contrast = c("TP_N1-TP_T2","TP_T2-TP_N1"), 
  group = c("TP_N1","TP_T2"))
```

TP2_vsTP4
```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'TP_T4-TP_T2','TP_T2-TP_T4'")
contrast_tbl = tibble(
  contrast = c("TP_T4-TP_T2","TP_T2-TP_T4"), 
  group = c("TP_T4","TP_T2"))
```


We subset the data to keep only cells from our timepoints that we want to compare first
```{r}
sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.05
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE
```

For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
```{r}
top_n_target = 250
# Then set multicore analysis to process faster
cores_system = 8
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type
```

Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
```{r}
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)
```


Perform MultiNicheNet’s cell-cell communication analysis
```{r}
multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```
Save output
```{r}
#SP_TP4_vs_TP2 <- multinichenet_output
SP_TPN1_vs_TP2 <- multinichenet_output
```

NOW TP_N1 AND TP_TP4

```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'TP_N1-TP_T4','TP_T4-TP_N1'")
contrast_tbl = tibble(
  contrast = c("TP_N1-TP_T4","TP_T4-TP_N1"), 
  group = c("TP_N1","TP_T4"))

```

We subset the data to keep only cells from our timepoints that we want to compare first
```{r}
sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.05
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE
```

For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
```{r}
top_n_target = 250
# Then set multicore analysis to process faster
cores_system = 8
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type
```

Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
```{r}
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)
```


Perform MultiNicheNet’s cell-cell communication analysis
```{r}
multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```
Save output
```{r}
SP_TPN1_vs_TP4 <- multinichenet_output
SP_TP2_vs_TP4 <- multinichenet_output
```


Barplots for Nichenet
```{r}
bm_t4_genes <- get_top_n_lr_pairs(BM_TPN1_vs_TP4$prioritization_tables, 50, rank_per_group = FALSE)
c(bm_t4_genes$ligand, bm_t4_genes$receptor)

bm_t2_genes <- get_top_n_lr_pairs(BM_TPN1_vs_TP2$prioritization_tables, 50, rank_per_group = FALSE)
c(bm_t2_genes$ligand, bm_t2_genes$receptor)

sp_t2_genes <- get_top_n_lr_pairs(SP_TPN1_vs_TP2$prioritization_tables, 50, rank_per_group = FALSE)
c(sp_t2_genes$ligand, bm_t2_genes$receptor)

sp_t4_genes <- get_top_n_lr_pairs(SP_TPN1_vs_TP4$prioritization_tables, 50, rank_per_group = FALSE)
c(sp_t4_genes$ligand, bm_t4_genes$receptor)
```


Creating my color scheme for nichnet
```{r}

mycols = c("B cells" = "#F28385", "CD4+ T cells" = "#C79ECF", "CD8+ T cells" = "#88D987", "DCs" = "#FED4A6", 
           "Macrophages" = "#6BAED6", "Monocytes" = "#F49C5D", "NK cells" = "#FFFF99", "Neutrophils" = "#F29FC6",
           "Pro-Erythroblasts" = "#76C1BB","Erythroblasts"= "#FF9999", "Neutrophil Progenitors" = "#A4D3E8",
           "Pre-B cells" = "#61B985",  
            "Pro-B cells" = "#AEBD93",
             "HSPCs"= "#C89A76")

nitchcols <- c("#F28385", "#C79ECF", "#88D987","#FED4A6", "#6BAED6", "#F49C5D", "#FFFF99", "#F29FC6")

# My cells of interest
#unique(BM_chabaudi.combined$cell_types)[-c(2,6,10)]

my_cells <- c("B.cells", "CD4..T.cells", "CD8..T.cells", "DCs", 
              "Macrophages", "Monocytes", "NK.cells", "Neutrophils" )

colors_sender = nitchcols %>% magrittr::set_names(my_cells)
colors_receiver = nitchcols %>% magrittr::set_names(my_cells)
            
```


Plot the output plots
```{r}
for (res_dta in c("BM_TPN1_vs_TP2_merged", "BM_TPN1_vs_TP4_merged", "SP_TPN1_vs_TP2", 
                  "SP_TPN1_vs_TP4", "SP_TP2_vs_TP4")) {
  multinichenet_output <- get(res_dta)
  prioritized_tbl_oi_all = get_top_n_lr_pairs(multinichenet_output$prioritization_tables, 50, rank_per_group = FALSE)
  prioritized_tbl_oi = multinichenet_output$prioritization_tables$group_prioritization_tbl %>%
  filter(id %in% prioritized_tbl_oi_all$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group) %>% left_join(prioritized_tbl_oi_all)
  prioritized_tbl_oi$prioritization_score[is.na(prioritized_tbl_oi$prioritization_score)] = 0
  
  senders_receivers = union(prioritized_tbl_oi$sender %>% unique(), prioritized_tbl_oi$receiver %>% unique()) %>% sort()
  
  #colors_sender = nitchcols %>% magrittr::set_names(senders_receivers) #Never to be used. my colors set instead
  #colors_receiver = nitchcols %>% magrittr::set_names(senders_receivers)
  
  par(cex=1.3,  # Adjusts text size globally
    cex.axis=1.3,  # Adjusts axis text size
    cex.lab=1.3 
    #font=2,
    #col="black"
    )
  
  options(repr.plot.width=30, repr.plot.height=20)
  circos_list = make_circos_group_comparison(prioritized_tbl_oi, colors_sender, colors_receiver)
  
  head(multinichenet_output$prioritization_tables)
  
  par(cex=1,  # Adjusts text size globally
    cex.axis=1,  # Adjusts axis text size
    cex.lab=1 
    #font=1,
    #col="black"
)
}
```

save the nichenet outputs
```{r}
nichenet_obj <- list(BM_TPN1_vs_TP2_merged, BM_TPN1_vs_TP4_merged, SP_TPN1_vs_TP2, 
                     SP_TPN1_vs_TP4, SP_TP2_vs_TP4)
names(nichenet_obj) <- c("BM_TPN1_vs_TP2_merged", "BM_TPN1_vs_TP4_merged", "SP_TPN1_vs_TP2", 
                         "SP_TPN1_vs_TP4", "SP_TP2_vs_TP4")

saveRDS(nichenet_obj,"/Users/joycekabagenyi/Documents/PhD_Project/Analysis/Nichnet/nichenet_objs_2024_v2.rds")
```



Visualization of the results of the cell-cell communication analysis

1. Circos plot of top-prioritized links
```{r}
# Looking at the top 50 predictions across all contrasts, senders, and receivers of interest
prioritized_tbl_oi_all = get_top_n_lr_pairs(multinichenet_output$prioritization_tables, 50, rank_per_group = FALSE)
```


```{r}
prioritized_tbl_oi = multinichenet_output$prioritization_tables$group_prioritization_tbl %>%
  filter(id %in% prioritized_tbl_oi_all$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group) %>% left_join(prioritized_tbl_oi_all)
prioritized_tbl_oi$prioritization_score[is.na(prioritized_tbl_oi$prioritization_score)] = 0

senders_receivers = union(prioritized_tbl_oi$sender %>% unique(), prioritized_tbl_oi$receiver %>% unique()) %>% sort()

colors_sender = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)
colors_receiver = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)

options(repr.plot.width=30, repr.plot.height=20)
circos_list = make_circos_group_comparison(prioritized_tbl_oi, colors_sender, colors_receiver)

head(multinichenet_output$prioritization_tables)
```


```{r}
senders_receivers = union(prioritized_tbl_oi$sender %>% unique(), prioritized_tbl_oi$receiver %>% unique()) %>% sort()

colors_sender = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)
colors_receiver = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)

options(repr.plot.width=30, repr.plot.height=20, cex = 1)
circos_list = make_circos_group_comparison(prioritized_tbl_oi, colors_sender, colors_receiver)

head(multinichenet_output$prioritization_tables)
```

****** Comparing bone marrow and spleen TP2 ******

Convert my seurat object into a singleR object # Starting with the bone marrow then spleen
```{r}
BM_chabaudi.combined <- subset(chabaudi_combined, TimePoint == "TP_T2") # Import RDS
BM_chabaudi.combined@meta.data[["consensuscall"]] <- ifelse(is.na(BM_chabaudi.combined@meta.data[["consensuscall"]]) == TRUE, BM_chabaudi.combined@meta.data[["Condition"]],BM_chabaudi.combined@meta.data[["consensuscall"]])
unique(BM_chabaudi.combined@meta.data[["consensuscall"]])

BM_chabaudi.combined$sampleID <- (ifelse(is.na(BM_chabaudi.combined@meta.data[["TimePoint"]]) == FALSE,
                                    paste(BM_chabaudi.combined@meta.data[["consensuscall"]],BM_chabaudi.combined@meta.data[["TimePoint"]], sep = "_"),BM_chabaudi.combined@meta.data[["Condition"]]))

sce_chabaudi = Seurat::as.SingleCellExperiment(BM_chabaudi.combined, assay = "RNA")
sce_chabaudi = alias_to_symbol_SCE(sce_chabaudi, "mouse") %>% makenames_SCE()
```

```{r}
sample_id = "sampleID"
group_id = "Tissue"
celltype_id = "cell_types"
covariates = NA
batches = NA
```



```{r}
SummarizedExperiment::colData(sce_chabaudi)$sampleID = SummarizedExperiment::colData(sce_chabaudi)$sampleID %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$TimePoint = SummarizedExperiment::colData(sce_chabaudi)$TimePoint %>% make.names()
SummarizedExperiment::colData(sce_chabaudi)$cell_types = SummarizedExperiment::colData(sce_chabaudi)$cell_types %>% make.names()

unique(colData(sce_chabaudi)$sampleID)
unique(colData(sce_chabaudi)$Tissue)
unique(colData(sce_chabaudi)$cell_types)

```
```{r}
senders_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
senders_oi = senders_oi[-c(7,8,10)] # Ignore erythroblasts and HSPCs
receivers_oi = SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %>% unique()
receivers_oi = receivers_oi[-c(7,8,10)] # Ignore erythroblasts and HSPCs
```

Subset or obeject if we want to use just a set of our cells
```{r}
sce_chabaudi = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,celltype_id] %in% c(senders_oi, receivers_oi)]

# cell type abundance and expression information from receiver and sender cell types
abundance_expression_info = get_abundance_expression_info(sce = sce_chabaudi, sample_id = "sampleID", group_id = "Tissue", celltype_id = "cell_types", min_cells = min_cells, senders_oi = senders_oi, receivers_oi = receivers_oi, lr_network = lr_network, batches = batches)

abundance_expression_info$abund_plot_sample
abundance_expression_info$abund_plot_group
```

```{r}
min_cells = 5 #Set minimum number of cells per cell type for any cell cluster to be considered for analysis
contrasts_oi = c("'spleen-bone_marrow','bone_marrow-spleen'")
contrast_tbl = tibble(
  contrast = c("spleen-bone_marrow","bone_marrow-spleen"), 
  group = c("spleen","bone_marrow"))

#"spleen"      "bone_marrow"
```

We subset the data to keep only cells from our timepoints that we want to compare first
```{r}
sce_chabaudi_subset = sce_chabaudi[, SummarizedExperiment::colData(sce_chabaudi)[,group_id] %in% contrast_tbl$group]
```

We then set cut offs for different parameters important for DGE analysis and what P-values to consider for further analysis
```{r}
logFC_threshold = 0.5 # Instead of 0.5
p_val_threshold = 0.05
fraction_cutoff = 0.05
# p_val_adj = TRUE 
p_val_adj = FALSE 
empirical_pval = FALSE
```

For the NicheNet ligand-target inference, we also need to select which top n of the predicted target genes will be considered (here: top 250 targets per ligand).
```{r}
top_n_target = 250
# Then set multicore analysis to process faster
cores_system = 8
n.cores = min(cores_system, union(senders_oi, receivers_oi) %>% length()) # use one core per receiver cell type
```

Define the weights of the prioritization of both expression, differential expression and NicheNet activity information
```{r}
prioritizing_weights_DE = c("de_ligand" = 1,
                         "de_receptor" = 1)
prioritizing_weights_activity = c("activity_scaled" = 2)

prioritizing_weights_expression_specificity = c("exprs_ligand" = 2,
                         "exprs_receptor" = 2)

prioritizing_weights_expression_sufficiency = c("frac_exprs_ligand_receptor" = 1)

prioritizing_weights_relative_abundance = c( "abund_sender" = 0,
                         "abund_receiver" = 0)

prioritizing_weights = c(prioritizing_weights_DE, 
                         prioritizing_weights_activity, 
                         prioritizing_weights_expression_specificity,
                         prioritizing_weights_expression_sufficiency, 
                         prioritizing_weights_relative_abundance)
```


Perform MultiNicheNet’s cell-cell communication analysis
```{r}
multinichenet_output = multi_nichenet_analysis(sce = sce_chabaudi_subset, celltype_id = celltype_id, sample_id = sample_id, group_id = group_id, 
                                lr_network = lr_network, ligand_target_matrix = ligand_target_matrix, contrasts_oi = contrasts_oi, contrast_tbl = contrast_tbl, batches = batches, covariates = covariates,
                                prioritizing_weights = prioritizing_weights, min_cells = min_cells, logFC_threshold = logFC_threshold, p_val_threshold = p_val_threshold,  
                                fraction_cutoff = fraction_cutoff, p_val_adj = p_val_adj, empirical_pval = empirical_pval, top_n_target = top_n_target, n.cores = n.cores, sender_receiver_separate = FALSE, verbose = TRUE)

```

```{r}
BM_vs_SP_T2 = prioritized_tbl_oi_all
```

```{r}
prioritized_tbl_oi_all = get_top_n_lr_pairs(multinichenet_output$prioritization_tables, 50, rank_per_group = FALSE)

prioritized_tbl_oi = multinichenet_output$prioritization_tables$group_prioritization_tbl %>%
  filter(id %in% prioritized_tbl_oi_all$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group) %>% left_join(prioritized_tbl_oi_all)
prioritized_tbl_oi$prioritization_score[is.na(prioritized_tbl_oi$prioritization_score)] = 0

senders_receivers = union(prioritized_tbl_oi$sender %>% unique(), prioritized_tbl_oi$receiver %>% unique()) %>% sort()

colors_sender = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)
colors_receiver = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)

options(repr.plot.width=30, repr.plot.height=20)
circos_list = make_circos_group_comparison(prioritized_tbl_oi, colors_sender, colors_receiver)

head(multinichenet_output$prioritization_tables)
```


